Apropos Replacement \newline (Initial Outline)

### 2011-12-26 paper.tex
Grammatical fixes

\documentclass[titlepage, a4paper, 12pt]{article}
\usepackage{listings}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[colorlinks=false]{hyperref}
\lstset{ %
language=sh,
basicstyle=\footnotesize,
numbers=none,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname
}
\title{Apropos Replacement \newline
(Initial Outline)}
\author{Abhinav Upadhyay $<$er.abhinav.upadhyay@gmail.com$>$ \and 
Joerg Sonnenberger $<$joerg@NetBSD.org$>$}
\begin{document}
\maketitle
\section{Abstract}
Unix like operating systems have withstood the test of time for 40 odd years for
several technical and philosophical reasons. One of those reasons is the
quality documentation they ship in the form of man pages. They are a single point
source of consultation for most of our routine work. But there has been a lack of
good search tool to accompany this documentation which would increase their
usefulness manifolds.

Traditionally \textit{apropos} [1] has been there to act as a search
interface but to a large extent it hasn't really lived up to it's expectation.
\textit{apropos} was
developed in the early days of Unix when computing resources were scarce and
which is the primary reason that it's design was kept so
minimal and simple.

In this paper the limitations of the traditional implementation of
\textit{apropos} are discussed along with brief details of how those limitations
have been solved in the implementation developed as part of this project and also
what other modern implementations are available out there.
\section{Introduction}
The classical version of \textit{apropos} has been implemented by simply
indexing the keywords in the NAME section of the man pages in a plain text file
(whatis.db) [2]
and performing searches on it. The reason for this simple design was most
probably lack of computing resources in the early days. The plain text file
consisting of keywords hardly takes few hundreds of kilo bytes of disk space
and performing search on a plain text file is also not very hard.

This simplified design of \textit{apropos(1)} resulted in limited utility and
usability.The searches are limited to the keywords in the NAME section of the
 keywords. If
only a user knows the exact keywords he gets the right results otherwise most of
the times the searches are full of bogus results or a dead end. In the modern
computing world where hard problems like searching the World Wide Web have been
solved [3] to sufficient degree, then it makes perfect sense that we leverage the advancement in technology to solve the problem of an effective man page
search tool, considering the fact that the man
pages consist of structured data, making the problem somewhat less complex.

In present times the machines are powerful and  capable of running the search
algorithms efficiently. Disk
space is also cheap enough that indexing of additional metadata from man pages
can be easily afforded. These are necessary prerequisites for building an
effective search tool.

@@

\subsection{Lack of support for free form queries}
The conventional \textit{apropos} is limited to the keywords used in the NAME
section
of the man pages. If the user specifies keywords relevant to the page which he
is looking but probably those keywords are not used in the NAME section,
then \textit{apropos} would not return
anything relevant. So in essence the user cannot use free form queries like
those supported by modern search applications.
\lstset{language = sh, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos ``add new user''
add new user: nothing appropriate

@@

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it has no
natural language support. For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query \textit{``make directories''} but it fails when the keyword \textit{``directory''} is
used in place of \textit{``directories''}, although the two words are based on
same root word.
\lstset{language = sh, caption={No support for stems or word with same roots},}
\begin{lstlisting}
$ apropos ``make directories''
mkdir (1) - make directories
$ apropos ``make directory''
make directory: nothing appropriate

@@

\subsection{Unintelligible Output}
Because of the way \textit{apropos} works, sometimes it's output can be
unintelligible and it can be very hard for the user to identify the results which
are relevant to him.
\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ \textit{apropos(1)} power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_class_display_register, pmf_system_suspend, pmf_system_resume,
pmf_system_shutdown, pmf_event_register, pmf_event_deregister,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
\subsection{Other Problems}
Apart from the search related problems there are a few issues related to
the way man pages are handled in NetBSD. The different aliases of the man pages
are stored on the filesystem in the from of hard (or soft) links and these
have to be mentioned in the makefiles explicitly using the MLINKS mechanism. This
approach works fine but it is a mess from maintenance point of view.
It should be possible to fix this by utilizing the index already built and
maintained by \textit{apropos} and related tools, but yet again, the
simplistic implementation of \textit{apropos} does not leave any room for
improvement.

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems as
part of this project.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the keywords defined in the NAME section of the man
pages, the keywords used throughout the body of the page also play a role.
For example queries like \textit{``installing new software package''}
will now get results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This enables support for more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

@@

and building an index of the data. \textit{libmandoc} [6]
and \textit{sqlite} [7] have been used for the two tasks respectively in this
implementation.

@@

interface to parse and build an AST (Abstract Syntax Tree) of the man page, 
and interface for traversing that tree in order to extract the data from nodes

@@

Due to space constraints it is not possibile to go into enough implementation details but in this section some important components of this project are
discussed in brief.

>>>>

\documentclass[titlepage, a4paper, 12pt]{article}
\usepackage{listings}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[colorlinks=false]{hyperref}
\lstset{ %
language=sh,
basicstyle=\footnotesize,
numbers=none,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname
}
\title{Apropos Replacement \newline
(Initial Outline)}
\author{Abhinav Upadhyay $<$er.abhinav.upadhyay@gmail.com$>$ \and 
Joerg Sonnenberger $<$joerg@NetBSD.org$>$}
\begin{document}
\maketitle
\section{Abstract}
Unix like operating systems have withstood the test of time for 40 odd years for
several technical and philosophical reasons. One of those reasons is the
quality documentation they ship in the form of man pages, which are a single
source of consultation for most of our routine work. However, so far, there has been a lack of
good search tool to accompany this documentation, which would increase their
usefulness manifolds.

Traditionally, \textit{apropos} [1] has been there to act as a search
interface but to a large extent it hasn't really lived up to it's expectation.
\textit{apropos} was
developed in the early days of Unix when computing resources were scarce and
that is the primary reason for its minimalistic and simple design.

In this paper, the limitations of the traditional implementation of
\textit{apropos} are discussed along with brief details of how these were overcome as part of this.
Additionally, we compare and contrast our approach with other modern implementations available out there.

\section{Introduction}
The classical version of \textit{apropos} has been implemented by simply
indexing the keywords in the NAME section of the man pages in a plain text file
(whatis.db) [2]
and performing searches on it. The reason for this simple design was most
probably lack of computing resources in the early days. The plain text file
consisting of keywords hardly takes few hundreds of kilo bytes of disk space
and performing search on a plain text file is quite easy.

This simplified design of \textit{apropos(1)} resulted in limited utility and
usability as the searches are limited to the keywords in the NAME section of the
 keywords. Only if
a user knows the exact keywords do they get the appropriate results; otherwise most of
the times the searches are full of irrelavant results or possibly a dead end. In the modern
computing world where hard problems like searching the World Wide Web have been
solved [3] to a sufficient degree, it makes perfect sense to leverage the advancement in technology 
in order to solve the problem. This problem is made even simpler considering that the man
pages consist of structured data, making the problem somewhat less complex.

In present times, the machines are powerful and  capable of running the search
algorithms efficiently. Disk
space is also sufficiently cheap that any extra space incurred by indexing of additional metadata from man pages
can be easily afforded. These are necessary prerequisites for building an
effective search tool.

@@

\subsection{Lack of support for free form queries}
The conventional \textit{apropos} is limited to the keywords used in the NAME
section
of the man pages. If the user specifies keywords relevant to the page which he
is looking but if those keywords are not used in the NAME section,
then \textit{apropos} would not return
anything relevant. So, in essence, the user cannot use free form queries like
those supported by modern search applications.
\lstset{language = sh, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos ``add new user''
add new user: nothing appropriate

@@

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it has no
natural language support. For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query \textit{``make directories''}, it fails when the keyword \textit{``directory''} is
used in place of \textit{``directories''} even though the two words are based on
same root word.
\lstset{language = sh, caption={No support for stems or word with same roots},}
\begin{lstlisting}
$ apropos ``make directories''
mkdir (1) - make directories
$ apropos ``make directory''
make directory: nothing appropriate

@@

\subsection{Unintelligible Output}
Because of the way \textit{apropos} works, sometimes it's output can be
unintelligible and it can be very hard for the user to identify relevant results.

\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ \textit{apropos(1)} power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_class_display_register, pmf_system_suspend, pmf_system_resume,
pmf_system_shutdown, pmf_event_register, pmf_event_deregister,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
\subsection{Other Problems}
Apart from the search related problems there are a few issues related to
the way man pages are handled in NetBSD. The different aliases of the man pages
are stored on the filesystem in the from of hard (or soft) links and these
have to be mentioned in the makefiles explicitly using the MLINKS mechanism. This
approach works fine but it is a mess from maintenance point of view.
It should be possible to fix this by utilizing the index already built and
maintained by \textit{apropos} and related tools, but yet again, the
simplistic implementation of \textit{apropos} does not leave any room for
improvement.

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems as
part of this project.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the keywords defined in the NAME section of the man
pages as the keywords used throughout the body of the page also play a role.
e.g. queries like \textit{``installing new software package''}
will now produce relevant results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation, the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This enables support for more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

@@

and building an index from the data. We have used \textit{libmandoc} [6]
and \textit{sqlite} [7] have for these the two tasks.

@@

interface to parse and build an AST (Abstract Syntax Tree) of the man page. It also provides
an interface for traversing that tree in order to extract the data from nodes

@@

Due to space constraints it is not possibile to go into enough implementation details, we discuss 
the most important components of this project  in the section below:
